---
title: "EDC101: Using Data Visualisation to Build Personal Brands"
subtitle: ""  
author: 
  - "Dr. Yiyang Gao"
date: '`r Sys.Date()`'
output:
  xaringan::moon_reader:
    css: ["xaringan-themer.css", "css/custom.css"]
    seal: false
    nature:
      slideNumberFormat: "%current%"
      highlightStyle: github
      highlightLines: true
      sealed: false
      ratio: 16:9
      self_contained: true
      countIncrementalSlides: true
    includes:
      after_body: header/insert-header-violet.html
---

class: middle, title
background-size: contain

<br><br>

# Using Data Visualisation to Build Personal Brands
#### EDC101 Guest Session

<br><br>

**Dr. Yiyang Gao**<br>
Sheffield Methods Institute, The University of Sheffield<br>
[y.gao@sheffield.ac.uk](mailto:y.gao@sheffield.ac.uk)

```{r setup, include=FALSE}
options(htmltools.dir.version = FALSE)


library(tidyverse)
library(xaringan)
library(xaringanExtra)
library(xaringanthemer)

# Defaults for code
knitr::opts_chunk$set(
  fig.width=9, fig.height=3.5, fig.retina=3,
  out.width = "100%",
  cache = FALSE,
  echo = TRUE,
  message = FALSE, 
  warning = FALSE,
  fig.show = TRUE,
  hiline = TRUE
)

# Global theme (keep close to SMI look-and-feel)
theme_set(
  theme(plot.background = element_rect(fill = "#F8F8F8", colour = "#F8F8F8"), 
        panel.background = element_rect(fill = "#F8F8F8", colour = "#F8F8F8"),
        legend.background = element_rect(fill = "#F8F8F8", colour = "#F8F8F8")
        )
  )
```

```{r xaringan-tile-view, echo=FALSE}
# Use tile overview by hitting the o key when presenting
xaringanExtra::use_tile_view()
```

```{r xaringan-logo, echo=FALSE}
# Add logo to top right (adjust path if needed)
xaringanExtra::use_logo(
  image_url = "header/smi-logo-white.png",
  exclude_class = c("inverse", "hide_logo"), 
  width = "180px", position = css_position(top = "1em", right = "2em")
)
```

```{r xaringan-themer, include=FALSE, warning=FALSE}
# University colours
primary_color <- "#131E29"
secondary_color <- "#440099"
tuos_blue <- "#9ADBE8"
white = "#F8F8F8"
tuos_yellow <- "#FCF281"
tuos_purple <- "#440099"
tuos_red <- "#E7004C"
tuos_midnight <- "#131E29"

style_uos <- style_duo_accent(
  primary_color = "#131E29",
  secondary_color = "#440099",
  colors = c(tuos_purple = "#440099", 
             grey = "#131E2960", 
             tuos_blue ="#9ADBE8",
             tuos_mint = "#00CE7C",
             ac_yellow = '#FEC740'),
  header_font_google = xaringanthemer::google_font("Source Serif Pro", "600", "600i"),
  text_font_google   = xaringanthemer::google_font("Source Sans Pro", "300", "300i", "600", "600i"),
  code_font_google   = xaringanthemer::google_font("Lucida Console"),
  header_h1_font_size = "2rem",
  header_h2_font_size = "1.5rem", 
  header_h3_font_size = "1.25rem", 
  text_font_size = "0.9rem",
  code_font_size = "0.65rem", 
  code_inline_font_size = "0.85rem",
  inverse_text_color = "#9ADBE8", 
  background_color = "#F8F8F8", 
  text_color = "#131E29", 
  link_color = "#005A8F", 
  inverse_link_color = "#F8F8F8",
  text_slide_number_color = "#44009970",
  table_row_even_background_color = "transparent", 
  table_border_color = "#44009970",
  text_bold_font_weight = 600
)

style_uos
```

```{r xaringan-panelset, echo=FALSE}
xaringanExtra::use_panelset()
```

```{r xaringanExtra, echo = FALSE}
xaringanExtra::use_progress_bar(color = "#F8F8F8", location = "top")
```

```{r xaringan-extra-styles, echo = FALSE}
xaringanExtra::use_extra_styles(
  hover_code_line = TRUE,
  mute_unhighlighted_code = TRUE
)
```

```{r share-again, echo=FALSE}
xaringanExtra::use_share_again()
```

```{r xaringanExtra-search, echo=FALSE}
xaringanExtra::use_search(show_icon = TRUE, auto_search = FALSE)
```

---

class: middle, inverse

# Sign in

---

class: middle

## Learning outcomes

.panel[.panel-name[What will I learn?]

By the end of this session you will be able to:

* Explain how data visualisation can support **personal branding** and employability.
* Identify **actionable steps** to build a website and GitHub portfolio.
* Shift mindset from *“I haven’t done it”* to *“I made a demo”* via small, reproducible projects.
* Use a short list of **free tools/resources** (Git/GitHub, R Markdown/Quarto, Shiny) to publish work online.
* Plan one **visible demo** you can complete this week.
]


---

class: inverse, middle

# Part I — Four Dimensions

---

### Four Dimensions of This Talk  

Action · Application · Mindset · Skills · Reflection

* **Action** — Build a website and maintain a portfolio through data visualisation.  
* **Application** — Examples that fit BA Digital Media and Society pathways.  
* **Mindset** — From “I haven’t done it” → “I made a demo”.  
* **Skills** — Tools and workflows for reproducible creativity.  
* **Reflection** — How dataviz underpins your personal brand.

---

class: inverse, middle

# Part II — Action

---

### Action (1): Build your personal website

* Calum’s website: <https://calumwebb.co.uk>  
* Calum’s GitHub repo: <https://github.com/cjrwebb/calumwebb>  
* Yiyang’s website: <https://yiyang-gao-1.github.io>  
* Yiyang’s GitHub repo: <https://github.com/yiyang-gao-1/yiyang-gao-1.github.io>  

**Takeaway:** Don’t *say* you can do it; **show** you can do it — move from static slides to interactive, reproducible sites.

---

### Action (2): Create your portfolio on GitHub

* EDC101 slides are built in **R Markdown** — version-controlled and reproducible on GitHub.  
* A strong GitHub portfolio collects projects, notes and materials — a transparent trace of learning.  
* Example module repo: <https://github.com/cjrwebb/smi105-teaching-materials/tree/main>.

**Why portfolio?** Not just a gallery. It’s evidence of practice and growth — *output drives input*.

---

### Action (3): Demonstrate capability

* Case study: **Sheffield post** — I didn’t know *MAIHDA* initially but had multilevel modelling experience.
* Read JD carefully → extract themes: **longitudinal, spatial, causal** MAIHDA; methods like **MCMC**; data: **Understanding Society**.
* Built a **demo project** to show readiness — it became key evidence in the application.

**Rule of thumb:** Don’t promise — **prototype**.

---

class: inverse, middle

# Part III — Application

---

### Turning learning into digital career opportunities

| Pathway | Example roles | Core skill sets | Demo ideas (data + storytelling) |
|---|---|---|---|
| Digital Media and Communications | Digital Media Manager; Social Media Account Manager; Digital Content Producer | Data visualisation; storytelling; content design; communication | Visualise social‑media trends (e.g., TikTok/X hashtag dynamics; compare engagement across campaigns) |
| UX / User Research | UX Designer; Usability Researcher; User Experience Analyst | Data analysis; visualisation; user‑centred testing | Map user journeys; create a mini dashboard of usability feedback |
| Marketing and Consumer Insights | Digital Marketing Analyst; Market Research Executive; Campaign Analyst | Data‑driven storytelling; visual reporting; consumer insight | Visualise campaign performance; analyse brand sentiment; build a consumer‑behaviour dashboard |
| Data Storytelling and Journalism | Data Journalist; Digital Storytelling Producer; Policy Communicator | Data narrative; interactive visualisation; ethical communication | Spotify listening‑trend story; misinformation network visual; map online discourse flows |
| Public Sector and NGOs | Communication Officer; Digital Engagement Lead; Policy Analyst | Public communication; social‑issue data presentation; impact evaluation | Protest movement maps |

**Tip:** You don’t need big data or complex models — **small, visible demos** (infographic, dashboard, short data story) published on GitHub + website.

---

class: inverse, middle

# Part IV — Mindset

---

### From “I haven’t done it” → “I made a demo”

* AI/LLMs reduce barriers — you can **prototype** quickly and learn by doing.  
* Success = **grit + self‑learning + proactive attitude**.  
* Expect confusion and mistakes → *draft → refine → repeat*.  
* **Output drives input:** creating demos is the most effective way to learn.

---

class: inverse, middle

# Part V — Skills and Handy Tools

---

### Skills and Tools — Self‑learning resources (all free)

**1. Git + GitHub (version control & portfolio)**  
* Track progress, collaborate, showcase work.  
* *Happy Git with R* — <https://happygitwithr.com/>

**2. RStudio + R Markdown / Quarto → HTML**  
* Turn analysis into shareable reports and web pages.  
* *R Markdown: The Definitive Guide* — <https://bookdown.org/yihui/rmarkdown/>  
* **Quarto** — <https://quarto.org/>

**3. R Shiny dashboards**  
* Build interactive visualisations and apps directly from R.  
* Shiny tutorial — <https://shiny.posit.co/r/>  
* Shiny gallery — <https://shiny.posit.co/gallery/>  
* Example (Calum): <https://webb.shinyapps.io/ipse/>

---

class: inverse, middle

# Part VI — Reflection

---

### Small actions, big momentum

* Pick **one** idea that excites you this week:  
  * Visualise your **Spotify** data  
  * Analyse a **TikTok** trend  
  * Map a **digital conversation**  
  * Recreate a chart from a favourite article  
* Document the process on **GitHub** → post on your **website / LinkedIn**  
* Reflect on what you learned

> Don’t wait until you’re ready. **Build, share, and learn out loud.**

